### 2024-10-25 analysis/preprocessing.qmd
checking the number of models and studies table 1
feat(preprocessing): average V and A accuracy classification
update encoding for alvarez to include validation model in mcc conversion. take average of valence and arousal classification for hu2022de, zhang2016br

```{r}
#| warning: false
library(stringr) 
# get metaMER df:
meta_df <- get_metaMER_df(path_2_studies = here::here('studies'))

@@

```{r}
#| echo: false
#| eval: true
library(stringr)
metaMER_results$studyREF <- substr(metaMER_results$citekey,1,nchar(metaMER_results$citekey)-2)
metaMER_results$studyREF <- str_replace_all(metaMER_results$studyREF,'([0-9]+)',' et al \\1')
metaMER_results$studyREF <- str_to_sentence(metaMER_results$studyREF)
```

@@

```{r}
#| eval: true
# Classify techniques according Hastie, Tibshirani, Friedman (2008)
# https://www.sas.upenn.edu/~fdiebold/NoHesitations/BookAdvanced.pdf
#
library(stringr)
metaMER_results$model_id <- tolower(metaMER_results$model_id)
metaMER_results$model_class_id <- 'Unclassified'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'lr|lm|pls|mlr|pcr|logistic regression|2d model full|pentagon|gaussian process regression|sparse bayesian regression|variational bayesian regression|logistic regression|lda|rda|regularized discriminant analysis|reguliarized discriminant analysis')] <- 'Linear Methods' # Class name from Elements of Stat.."
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'logistic regression|lda|rda|regularized discriminant analysis|reguliarized discriminant analysis')] <- 'Linear Classification'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'mars|gam')]<-'Additive Trees and Related Methods' #
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rbf|gmm|local|polynomial|polygonal|knn|mars|gam')]<-'Kernel Smoothing, Additive and KNN'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'adaboost|gradient')]<-'Boosting'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'nn|gru|lstm|ltsm|long short term memory|rprop|mcan')]<-'Neural Nets'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'svm|svr|support vector regression|smoreg|smo ')]<-'Flexible Discriminants'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'knn')]<-'Prototype methods'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rf|extremely randomized tree regression|random forest|adaboost|gradient')]<-'Random Forests'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rf')]<-'Ensemble Learning'

@@

cat(paste("\nWhere", length(unique(metaMER_results$citekey[metaMER_results$model_category=='classification'])), "are classification studies"))
# note that we have some classification studies that also do regression and vice versa?
# THIS IS CORRECT (updated 12 July 2024):
# [1] "We have 1238 observations"
# [1] "We have 37 studies"
# [1] "Where 23 are regression studies"
# [1] "Where 14 are classification studies"

@@

ez::ezDesign(data = R_studies, x = unique_id, y = citekey)

@@

```{r}
library(ggplot2)

@@

```{r}
library(tidyverse)
R_studies$citekey <- factor(R_studies$citekey)
R_studies$dimension <- factor(R_studies$dimension)

@@

library(ggrepel)

@@

library(ggdist)

@@

## Create descriptive table for the manuscript

@@

# take accuracy as MCC in those studies that have not specified mcc
C_studies$accuracy_converted_to_mcc <- 0
C_studies$accuracy_converted_to_mcc[C_studies$citekey=='bai2016di' | 
                  C_studies$citekey=='bai2017mu' | 
                  C_studies$citekey=='hizlisoy2021mu' |
                  C_studies$citekey=='hu2022de' |    
                  C_studies$citekey=='yeh2014po' |
                  C_studies$citekey=='zhang2016br' |
                  C_studies$citekey=='zhang2017fe' 
                  & C_studies$measure=='accuracy']<-1

C_studies$measure[C_studies$citekey=='bai2016di' | 
                  C_studies$citekey=='bai2017mu' | 
                  C_studies$citekey=='hizlisoy2021mu' |
                  C_studies$citekey=='hu2022de' |    
                  C_studies$citekey=='yeh2014po' |
                  C_studies$citekey=='zhang2016br' |
                  C_studies$citekey=='zhang2017fe' 
                  & C_studies$measure=='accuracy']<-'mcc'
table(C_studies$measure,C_studies$citekey)
C_studies <- dplyr::filter(C_studies, str_detect(measure,'mcc')) #  
dim(C_studies)
table(C_studies$measure,C_studies$accuracy_converted_to_mcc)
t.test(C_studies$values,y = C_studies$accuracy_converted_to_mcc) # There is a difference

```
#### Get the number of categories in the classification task

@@

```{r}
#| eval: false
library(ggplot2)

@@

```{r}
library(tidyverse)
C_studies$citekey <- factor(C_studies$citekey)

@@

```{r}
#| eval: true
tmp <- drop_na(C_studies)
library(ggdist)

@@

## Create descriptive table for the manuscript

>>>>

```{r}
#| warning: false
library(stringr)
library(tidyverse)
library(ggrepel)
library(ez)
library(ggdist)

# get metaMER df:
meta_df <- get_metaMER_df(path_2_studies = here::here('studies'))

@@

```{r}
#| echo: false
#| eval: true
metaMER_results$studyREF <- substr(metaMER_results$citekey,1,nchar(metaMER_results$citekey)-2)
metaMER_results$studyREF <- str_replace_all(metaMER_results$studyREF,'([0-9]+)',' et al \\1')
metaMER_results$studyREF <- str_to_sentence(metaMER_results$studyREF)
```

@@

```{r}
#| eval: true
# Classify techniques according Hastie, Tibshirani, Friedman (2008)
# https://www.sas.upenn.edu/~fdiebold/NoHesitations/BookAdvanced.pdf
#
metaMER_results$model_id <- tolower(metaMER_results$model_id)
metaMER_results$model_class_id <- 'Unclassified'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'lr|lm|pls|mlr|pcr|logistic regression|2d model full|pentagon|gaussian process regression|sparse bayesian regression|variational bayesian regression|logistic regression|lda|rda|regularized discriminant analysis|reguliarized discriminant analysis')] <- 'Linear Methods' # Class name from Elements of Stat.."
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'logistic regression|lda|rda|regularized discriminant analysis|reguliarized discriminant analysis')] <- 'Linear Classification'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'mars|gam')]<-'Additive Trees and Related Methods' #
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rbf|gmm|local|polynomial|polygonal|knn|mars|gam')]<-'Kernel Smoothing, Additive and KNN'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'adaboost|gradient')]<-'Boosting'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'nn|gru|lstm|ltsm|long short term memory|rprop|mcan')]<-'Neural Nets'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'svm|svr|support vector regression|smoreg|smo ')]<-'Flexible Discriminants'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'knn')]<-'Prototype methods'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rf|extremely randomized tree regression|random forest|adaboost|gradient')]<-'Random Forests'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rf')]<-'Ensemble Learning'

@@

cat(paste("\nWhere", length(unique(metaMER_results$citekey[metaMER_results$model_category=='classification'])), "are classification studies"))
# note that we have some classification studies that also do regression and vice versa?
# THIS IS CORRECT (updated 24 October 2024):
# [1] "We have 1284 observations"
# [1] "We have 36 studies"
# [1] "Where 22 are regression studies"
# [1] "Where 14 are classification studies"

@@

ez::ezDesign(data = R_studies, x = unique_id, y = citekey)+
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank())

@@

@@

```{r}
R_studies$citekey <- factor(R_studies$citekey)
R_studies$dimension <- factor(R_studies$dimension)

@@

@@

@@

## Create descriptive table for the manuscript  (Table 1, column 1)

@@

# take accuracy as MCC in those studies that have not specified mcc
C_studies$accuracy_converted_to_mcc <- 0
C_studies$accuracy_converted_to_mcc[
                  C_studies$unique_id==
'alvarez2023ri-spotify-randomforest-spotify-acousticBrainz-modelvalidation'|
                  C_studies$citekey=='bai2016di' | 
                  C_studies$citekey=='bai2017mu' | 
                  C_studies$citekey=='hizlisoy2021mu' |
                  C_studies$citekey=='hu2022de' |    
                  C_studies$citekey=='yeh2014po' |
                  C_studies$citekey=='zhang2016br' |
                  C_studies$citekey=='zhang2017fe' 
                  & C_studies$measure=='accuracy']<-1

C_studies$measure[C_studies$unique_id==
'alvarez2023ri-spotify-randomforest-spotify-acousticBrainz-modelvalidation'|
                  C_studies$citekey=='bai2016di' | 
                  C_studies$citekey=='bai2017mu' | 
                  C_studies$citekey=='hizlisoy2021mu' |
                  C_studies$citekey=='hu2022de' |    
                  C_studies$citekey=='yeh2014po' |
                  C_studies$citekey=='zhang2016br' |
                  C_studies$citekey=='zhang2017fe' 
                  & C_studies$measure=='accuracy']<-'mcc'
table(C_studies$measure,C_studies$citekey)
C_studies <- dplyr::filter(C_studies, str_detect(measure,'mcc')) #  
dim(C_studies)

C_studies <- C_studies[!is.na(C_studies$values),]

table(C_studies$measure,C_studies$accuracy_converted_to_mcc)
t.test(C_studies$values,y = C_studies$accuracy_converted_to_mcc) # There is a difference

# Average across valence and arousal for VA classification

length(unique(paste0(C_studies$unique_id, C_studies$dimension)))

ez::ezDesign(data = C_studies, x = unique_id, y = citekey)+
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank())

```{r}
va_studies <- c("zhang2016br", "hu2022de")

va_acc <- C_studies |> 
  filter(citekey %in% va_studies) |> 
  group_by(unique_id, measure, statistic) |> 
  mutate(values = mean(values), 
         accuracy_converted_to_mcc = mean(accuracy_converted_to_mcc))

va_acc$dimension <- "classification"

va_acc <- va_acc |> distinct()

C_studies <- C_studies[!C_studies$citekey %in% va_studies,]

C_studies <- rbind(C_studies, va_acc)

### Revisualize studies
```{r}

length(unique(paste0(C_studies$unique_id, C_studies$dimension)))

ez::ezDesign(data = C_studies, x = unique_id, y = citekey)+
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank())

#### Get the number of categories in the classification task

@@

```{r}
#| eval: false

@@

```{r}
C_studies$citekey <- factor(C_studies$citekey)

@@

```{r}
#| eval: true
tmp <- drop_na(C_studies)

@@

## Create descriptive table for the manuscript (Table 1, column 2)

CA: I think this should be correct, though may need to update after averaging across VA for some classification studies (zhang2016, hu2022)
